---
title: "Chapter 5 Practice"
author: "Micah Elias"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Downloads 
```{r}
install.packages("nycflights13")
```


5.1 Prerequisites 
```{r cars}
library("nycflights13")
library("tidyverse")
```

5.1.2 nycflights13
```{r}
?nycflights13
flights
view(flights)
```

5.2 dplyr basics 

```{r}
#dplyr functions wont change structure of the input, unless saves at variable
filter(flights, month == 1, day ==1)
jan1 <- filter(flights, month ==1, day ==1)

# wrapping a variable creation command in () will print results 
(dec25 <- filter(flights, month == 12, day == 25))
```

5.2.1 Comparisons 
```{r}
## == is needed to be equal (not =)
filter(flights, month == 1)
```

Logical Operators 
```{r}
#boolean operation, each criteria needs its own T/ F statement
filter(flights, month == 11 | month == 12)

# shorthand x%in% y
nov_dec <- filter(flights, month %in% c(11,12))
nov_dec

# so, !x | !y is the same as !(x&y)

#these two are the same 
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 12, dep_delay <= 120)

```


5.2.4 Exercises 
1.1. flights that had an arrival delay of two or more hours 
```{r}
filter(flights, arr_delay >= 60)
```

1.2. flew to houston 
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

1.3. operated by american, united, delta 
```{r}
filter(flights, carrier == "UA" | carrier == "AA" | carrier == "DL")

```

1.4. departed in summer
```{r}
filter(flights, month == 7 | month == 8 | month == 9)
```

1.5. arrived more than 2 hours late but didnt leave late 
```{r}
filter(flights, arr_delay >= 120 & dep_delay <= 0)
```

1.7 between midnight and 6am 
```{r}
filter(flights, dep_time >= 1 & dep_time <= 600)
```

2. the between() command 
```{r}
?between
filter(flights, between(dep_time, 1, 600))
```

3. missing dep_time
```{r}
filter(flights, dep_time == NA)
```

5.3 Arange Rows
```{r}
arrange(flights, year, month, day)
arrange(flights, desc(dep_delay))
```

5.3.1 Exercises
1 arrange to sort 
```{r}
arrange(flights, desc(dep_time == is.na(TRUE)))
```

2. most delayed 
```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_time)
```

5.5 Mutate 
```{r}
flights_sml <- 
  select(flights, 
        year:day,
        ends_with("delay"),
        distance,
        air_time
        )

mutate(flights_sml,
       gain = dep_delay - arr_delay, 
       speed = distance / air_time * 60,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )


```

Only keep new variable
```{r}
transmute(flights, 
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain/ hours)
```

